---
title: "Report Iraceplot Package"
output: pdf_document
always_allow_html: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include= FALSE}
library(iraceplot)
```

# General information

The provided execution of irace statistics are: 

- Iterations: `r {length(iraceResults$allElites)}`
- Configurations: `r {nrow(iraceResults$allConfigurations)}`
- Instances: `r {nrow(iraceResults$experiments)}`
- Experiments: `r {nrow(iraceResults$experimentLog)}`
- Elite configurations: `r {length(iraceResults$allElites[[length(iraceResults$allElites)]])}`

## Elite configurations
The final best configurations found by irace are:

```{r,fig.align="center", echo=FALSE}
knitr::kable(iraceResults$allConfigurations[iraceResults$allElites[[length(iraceResults$allElites)]],])
```

The ``parallel_coord`` function creates a parallel coordinate plot of a set of selected configurations. 

```{r,fig.align="center", fig.width= 7, echo=FALSE}
parallel_coord(iraceResults)
```

## Parameter values sampling
The frequency of the parameter values sampled by irace may indicate high performing areas:

```{r,fig.align="center", fig.width= 7, echo=FALSE}
sampling_frequency(iraceResults)
```

## Testing performance

The test performance allows to evaluate if the configurations found by irace. 

The final elite configuration performance on the test set:

```{r,fig.align="center", fig.width= 7, echo=FALSE}
if (iraceResults$scenario$testInstances == "" || is.null(iraceResults$scenario$testInstances)) {
  cat("No test performed.\n")
} else {
  boxplot_test(iraceResults, type = "best")
}
```

Iteration elite configuration performance on the test set:

```{r,fig.align="center", fig.width= 7, echo=FALSE}
if (iraceResults$scenario$testIterationElites) {
   boxplot_test(iraceResults, type = "all")
} else{
  cat("No iteration test performed.\n")
}
```

## Training performance 

Final elite configurations performance on the training set. 

```{r,fig.align="center", fig.width= 7}
boxplot_training(iraceResults)
```