---
title: "Report generated by the iraceplot package"
output: 
  html_document:
    toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include= FALSE}
library(iraceplot)
library(tidyr, quietly = TRUE)
library(dplyr, quietly = TRUE)
niterations <- length(iraceResults$allElites)
ninstances <- nrow(iraceResults$experiments)
irace_log <- iraceResults$experimentLog
byiterations <- NULL
for (i in 1:niterations) {
  nconfig <- length(unique(irace_log[irace_log[,"iteration"]==i,"configuration"]))
  nbest <- length(iraceResults$allElites[[i]])
  ninstances <- length(unique(irace_log[irace_log[,"iteration"]==i,"instance"]))
  nexperiments <- nrow(irace_log[irace_log[,"iteration"]==i,])
  bestid <- iraceResults$iterationElites[[i]]
  byiterations <- rbind(byiterations, c(i, nconfig, ninstances, nexperiments, nbest, bestid))
}
colnames(byiterations) <- c("iteration", "configurations", "instances", "experiments", "elites", "best id")

byinstance <- NULL
instances <- iraceResults$scenario$instances[iraceResults$state$.irace$instancesList[1:ninstances, "instance"]]
for (i in 1:length(instances)) {
  iname <- basename(instances[i])
  nexperiments <-  nrow(irace_log[irace_log[,"instance"]==i,])
  imean <- mean(iraceResults$experiments[i,], na.rm=TRUE)
  ibest <- min(iraceResults$experiments[i,], na.rm=TRUE)
  iworst <- max(iraceResults$experiments[i,], na.rm=TRUE)
  imedian <- median(iraceResults$experiments[1,], na.rm = TRUE)
  bestid <- which.min(iraceResults$experiments[i,])
  byinstance <- rbind(byinstance, c(iname, nexperiments, imean, imedian, ibest, iworst, bestid))
}
colnames(byinstance) <- c("instance", "experiments", "mean", "median", "best", "worst", "best id" )
best_elites <- as.character(iraceResults$allElites[[length(iraceResults$allElites)]])
```

# General information

- Iterations: `r {niterations}`
- Configurations: `r {nrow(iraceResults$allConfigurations)}`
- Instances: `r {ninstances}`
- Experiments: `r {nrow(iraceResults$experimentLog)}`
- Elite configurations: `r {length(iraceResults$allElites[[niterations]])}`

## By iteration

```{r,fig.align="center", echo=FALSE}
knitr::kable(byiterations)
```

## By instance

```{r,fig.align="center", echo=FALSE}
knitr::kable(byinstance)
```





# Parameter settings
## Elite configurations

The final best configurations found by irace are:

```{r,fig.align="center", echo=FALSE}
knitr::kable(iraceResults$allConfigurations[best_elites, , drop=FALSE],
             row.names = FALSE)
```

## Parallel coordinates visualization (only elites)

```{r,fig.align="center", fig.width= 7, echo=FALSE}
parallel_coord(iraceResults)
```

## Parameter values sampling
The frequency of the parameter values sampled by irace:

```{r,fig.align="center", fig.width= 7, echo=FALSE}
sampling_frequency(iraceResults)
```



# Performance

## Testing performance

Performance of the final elite configurations on the test instances:

```{r,fig.align="center", echo=FALSE}
if (iraceResults$scenario$testInstances == "" || is.null(iraceResults$scenario$testInstances)) {
  cat("No test instances given.\n")
} else {
best_results <- as_tibble(iraceResults$testing$experiments[, best_elites])
best_results <- best_results %>% pivot_longer(all_of(best_elites), names_to="ID") %>% group_by(ID) %>% summarise(n_instances = sum(!is.na(value)), mean = mean(value), sd = sd(value), median = median(value), min = min(value), max = max(value))
best_results <- best_results[match(best_elites, best_results$ID), , drop = FALSE]
knitr::kable(best_results, row.names = FALSE)
}
```

The final elite configuration performance on the test set:

```{r,fig.align="center", fig.width= 7, echo=FALSE}
if (iraceResults$scenario$testInstances == "" || is.null(iraceResults$scenario$testInstances)) {
  cat("No test instances given.\n")
} else {
  boxplot_test(iraceResults, type = "best")
}
```

Iteration elite configuration performance on the test set:

```{r,fig.align="center", fig.width= 7, echo=FALSE}
if (iraceResults$scenario$testIterationElites) {
   boxplot_test(iraceResults, type = "all", show_points=FALSE)
} else{
  cat("Iteration elites were not tested.\n")
}
```

## Training performance 

Performance of the final elite configurations on the training instances:

```{r,fig.align="center", echo=FALSE}
best_results <- as_tibble(iraceResults$experiments[, best_elites])
best_results <- best_results %>% pivot_longer(all_of(best_elites), names_to="ID") %>% group_by(ID) %>% summarise(n_instances = sum(!is.na(value)), mean = mean(value), sd = sd(value), median = median(value), min = min(value), max = max(value))
best_results <- best_results[match(best_elites, best_results$ID), , drop = FALSE]
knitr::kable(best_results, row.names = FALSE)
```

Final elite configurations performance on the training set:

```{r,fig.align="center", fig.width= 7, echo=FALSE}
boxplot_training(iraceResults)
```

# Races overview
```{r,fig.align="center", fig.width= 7, echo=FALSE}
plot_experiments_matrix(iraceResults)
```


