---
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{iraceplot_package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<h1 style="text-align:center">Iraceplot Package</h1>

<h3>Introduction</h3>

<p>
  Escribir introduccion
</p>


<h3>Functions</h3>

<p>
  Escribir sobre las funciones
</p>

<div>
  <ul>
    <li>iboxplot_test()</li>
    <li>iscatter_test()</li>
    <li>iparcoord()</li>
    <li>iparallelcoord()</li>
    <li>iparameter_freq()</li>
    <li>iparameter_frequency_iteration()</li>
    <li>iparameter_density_iteration()</li>
    <li>iparameter_compare_iteration()</li>
    <li>iheatmap()</li>
    <li>ibp()</li>
    <li>iscatter()</li>
  </ul>
</div>

<h3>Install Iraceplot</h3>

For installing iraceplot you need to install the devtools package:
```
install.packages("devtools")
```

Currently, the Irace Studio package can be installed from Gtihub:

```
devtools::install_github("pabloOnate/iraceplot")
```
<h3>How to use</h3>

This is a basic example which shows you how to solve a common problem:
```{r setup}
library(iraceplot)
```
you need to have the file generated by irace de rdata loaded, you must replace the path of the file with yours
```{r}
load("../data/iraceResults.rda")
```
<h3>Testing</h3>

<h4>Function iboxplot_test()</h4>

<p>

</p>

<p>
  Cantidatos elite de la ultima iteracion
</p>

<div style="text-align: center">
```{r, fig.width= 7}
iboxplot_test(iraceResults, type = "last", distance_min = TRUE)
iboxplot_test(iraceResults, type = "last")
```
</div>

<p>
  Candidatos elite con todas las iteraciones
</p>

<div style="text-align: center">
```{r, fig.width= 7}
iboxplot_test(iraceResults, distance_min = TRUE)
iboxplot_test(iraceResults)
```
</div>

<p>
  Candidatos elite de las mejores iteraciones
</p>

<div style="text-align: center">
```{r, fig.width= 7}
iboxplot_test(iraceResults, type = "best", distance_min = TRUE)
iboxplot_test(iraceResults, type = "best")
```
</div>

<h4>Function iscatter_test()</h4>

<p>
  Nos permite comparar dos configuraciones elites
</p>

<p>
  un scatter plot de las configuraciones ingresadas
</p>

<div style="text-align: center">
```{r, fig.width= 7}
iscatter_test(iraceResults, idVector = c("92","119"), distance_min = TRUE)
iscatter_test(iraceResults, idVector = c("92","119"))
```
</div>


<h4>Function iparallelcoord()</h4>

<p>

</p>

<div style = "text-align:center; width=50px">
```{r, fig.width=5}
iparallelcoord(iraceResults)
iparallelcoord(iraceResults,idConfiguration = c(20,50,100,300,500,600,700))
iparallelcoord(iraceResults, param_names = c("algorithm","alpha","rho","q0","rasrank"))
iparallelcoord(iraceResults,idConfiguration = c(20,50,100,300,500,600,700), param_names =  c("algorithm","alpha","rho","q0","rasrank"))
```
</div>

<h4>Function iparameter_freq</h4>

```{r, fig.height= 7, fig.width=7}
iparameter_freq(iraceResults)
iparameter_freq(iraceResults,param_names = c("algorithm","alpha","rho","q0","rasrank"))
```

<h4>Function iparamater_frequency_iteration()</h4>

```{r, fig.width= 7}
iparameter_frequency_iteration(iraceResults,parameter = "algorithm")
iparameter_frequency_iteration(iraceResults,parameter = "alpha")
```

<h4>Function iparamater_density_iteration()</h4>

```{r, fig.width= 7}
iparameter_density_iteration(iraceResults,parameter = "alpha")
```

<h4>Function iparamater_compare_iteration()</h4>

```{r, fig.width= 7}
iparameter_compare_iteration(iraceResults, parameter = "alpha")
```

<h4>Function iheatmap()</h4>

```{r, fig.width= 7}
iheatmap(iraceResults)
```

<h3>Training</h3>

<h4>Function ibp()</h4>

```{r, fig.width= 7}
ibp(iraceResults)
ibp(iraceResults,distance_min = TRUE)
ibp(iraceResults,numberIteration = c(100,400,700,710), distance_min = TRUE)
```

<h4>Function iscatter()</h4>

```{r, fig.width= 7}
iscatter(iraceResults, idVector = c(700,720))
iscatter(iraceResults, idVector = c(700,720), distance_min = TRUE)
```
