---
title: "Quick Start"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE}
#library(iraceplot)
load("../data/iraceResults.rda")
```

This guide aims at showing different examples of how the data produced by the irace configurator
can be visualized to obtain information about algorithm performance or high-performing parameters 
settings.

## About the execution of irace

To use the methods provided by this package you must have an irace data object, 
this object is saved as an Rdata file (irace.Rdata by default) after each irace 
execution. 

During the configuration procedure irace evaluates several candidate configurations 
(parameter settings) on different training insrances, creating an algorithm performance 
data set we call the *training data set*. This information is thus, the data
that irace had access to when configuring the algorithm. 

You can also enable the test evaluation option in irace, in which a set of elite 
configurations will be evaluated on a set of test instances after the execution of 
irace is finished. Nota that this option is not enabled by default and you
must provide the test instances in order to enable it. The performance obtained 
in this evalaution is called the *test data set*. This evaluation helps assess
the results of the configuration in a more "real" setup. For example, we can assess 
if the configuration process incurred in overtuning or if a type of instance
was underrepresented in the training set. We note that irace allows to perform
the test evaluations to the final elite configurations and to the elite configurations
of each iterations. For information about the irace setup we refer you to the irace 
package user guide. 

Note: Before executing irace, consider setting the test evaluation option of irace. 

Once irace is executed, you can load the irace data object in the R console by:

```{r eval=FALSE} 
load("path/to/irace.Rdata")
```

This will load in the console an object called `iraceResults` which contains all data
generated and used by the configurator.

## Configurations

Once irace is executed, the first thing you might want to do is to visualize how the
best configurations look like. You can do this with the `parallel_coord` method.

```{r fig.align="center"}
parallel_coord(iraceResults)
```

This plot shows the final elite configurations settings, each line represents one 
configuration. Note that the number of iterations is shown in the scale on the right
and the final elite configurations belong to the last iteration. To visualize 
configurations  considered as elite in each iteration use the `iterations` option:

```{r fig.align="center"}
parallel_coord(iraceResults, iterations=1:iraceResults$state$nbIterations)
```

You can also visualize all configurations sampled in one or more iterations disabling the 
`only_elite` option:

```{r fig.align="center"}
parallel_coord(iraceResults, iterations=4, only_elite=FALSE)
```


