---
title: "The iraceplot package: User Guide"
output: 
  html_document:
    toc: true
---

```{r echo=FALSE, prompt=FALSE, message=FALSE, warning=FALSE}

library(iraceplot, quietly = TRUE)
library(plotly, quietly = TRUE)
iraceResults <- read_logfile(system.file(package="iraceplot", "exdata", "guide-example.Rdata", mustWork = TRUE))
```

## Introduction


The iraceplot package provides a set of functions to create plots to visualize 
the configuration data generated by the configuration process implemented in
the irace package. 

The configuration process performed by irace will show ar the end of the execution one
or more configurations that are the best performing configurations found. This package
provides a set of functions that allow to further assess the performance of these
configurations and provides support to obtain insights about the details of the configuration
process.


# Installation
## Install iraceplot from CRAN
```
install.packages("iraceplot")
```

## Install iraceplot from github

For installing iraceplot you need to install the devtools package:
```
install.packages("devtools")
devtools::install_github("auto-optimization/iraceplot/")
```

## Executing irace

To use the methods provided by this package you must have an irace data object, 
this object is saved as an Rdata file (irace.Rdata by default) after each irace 
execution. 

During the configuration procedure irace evaluates several candidate configurations 
(parameter settings) on different training insrances, creating an algorithm performance 
data set we call the *training data set*. This information is thus, the data
that irace had access to when configuring the algorithm. 

You can also enable the test evaluation option in irace, in which a set of elite 
configurations will be evaluated on a set of test instances after the execution of 
irace is finished. Nota that this option is not enabled by default and you
must provide the test instances in order to enable it. The performance obtained 
in this evalaution is called the *test data set*. This evaluation helps assess
the results of the configuration in a more "real" setup. For example, we can assess 
if the configuration process incurred in overtuning or if a type of instance
was underrepresented in the training set. We note that irace allows to perform
the test evaluations to the final elite configurations and to the elite configurations
of each iterations. For information about the irace setup we refer you to the irace 
package user guide. 

Note: Before executing `irace`, consider enabling testing in irace. 

Once irace is executed, you can load the irace log in the R console as previously 
shown.


## How to use

This is a basic example that shows how to use `iraceplot`:

```{r eval=FALSE}
library(iraceplot)
```

To use the functions its required to have the log file generated by `irace` (`logFile` 
option in the `irace` package). This file is commonly saved in the in the directory 
in which `irace` was executed with the name `irace.Rdata`.

# Load  irace log data

You can load the data from the log file as follows:
```{r eval=FALSE}
iraceResults <- irace::read_logfile("~/path/to/irace.Rdata")
```

For more details about the contents of the log file, please go to the [user guide](https://cran.r-project.org/web/packages/irace/vignettes/irace-package.pdf) of the irace package.

However, most functions directly accept a filename and do not require you to load the data yourself. We only mention here for completeness.

# Visualizing irace configuration data
In the following, we provide an example how the functions implemented in this package
can be used to visualize the information generated by irace.


## Configurations

Once irace is executed, the first thing you might want to do is to visualize how the
best configurations look like. You can do this with the `parallel_coord` method:

```{r fig.align="center", fig.width= 7, message=FALSE, prompt=FALSE}
parallel_coord(iraceResults)
```

The plot shows by default the final elite configurations (last iteration), each 
line in the plot represents one configuration. The plot produced by `parallel_coord` 
can help you to both the distribution of the parameter values of a set of configurations 
and the common associations between these values. By default, the plot colors the 
lines using the number of instances evaluated by the configuration. You can use 
the `color_by_instances` argument to choose between coloring the lines by the 
number of instances executed or the last iteration the configurationwas executed on. 

To visualize the configurations considered as elites in each iteration use the 
`iterations` option. You can select one or more iterations. For example, to select
all iterations executed by irace:

```{r fig.align="center", fig.width= 7, message=FALSE, prompt=FALSE}
parallel_coord(iraceResults, iterations=1:iraceResults$state$nbIterations, 
               color_by_instances = FALSE)
```

You can also visualize all configurations sampled in one or more iterations disabling the 
`only_elite` option. For example, to visualize configurations sampled in iterations 1 to 9:

```{r fig.align="center", fig.width= 7, message=FALSE, prompt=FALSE}
parallel_coord(iraceResults, iterations=1:9, only_elite=FALSE)
```

If you are looking for something more flexible and you would like to provide
your own set of configurations, you can use the `plot_configurations()` function. 
This function generates a parallel coordinates plot (similar to the ones generated
by `parallel_coord()`) when provided with an arbitrary set of configurations and a 
parameter space object. The configurations must be provided in the format
in which irace handles configurations: a dataframe with parameter in each column.
For information about these formats, please chech the irace package [user guide](https://cran.r-project.org/web/packages/irace/vignettes/irace-package.pdf).
As an example, this lines display all elite configurations:

```{r fig.align="center", fig.width= 7, message=FALSE, prompt=FALSE}
all_elite <- iraceResults$allConfigurations[unlist(iraceResults$allElites),]
parameters <- iraceResults$parameters
plot_configurations(all_elite, parameters)
```

The `parallel_cat` function displays all configurations sampled in a set
of iterations. The plot groups parameter values in intervals and thus, it can be 
useful to visualize more tendencies of association between parameters values. As 
in the previous functions you can use the `iterations` argument to select the 
iterations from which configurations should be selected. For example, to visualize 
configurations sampled on iterations 3, 4 and 5:

```{r fig.align="center", fig.width= 7, fig.height=7, message=FALSE, prompt=FALSE}
parallel_cat(irace_results = iraceResults, iterations = c(3,4,5) )
```

You can also adjust the number of value intervals that are displayed for each parameter
using the `n_bins` parameter which by default is set to 3. Also, it is possible to
select a subset of configurations using the `id_configurations` argument providing 
a vector a configuration ids.

Both in the `parallel_coord` and `parallel_cat` functions parameters can be selected 
using the `param_names` argument. For example to select parameters un `parallel_cat`:

```{r fig.align="center", fig.width= 7, fig.height=6, message=FALSE, prompt=FALSE}
parallel_cat(irace_results = iraceResults, 
             param_names=c("algorithm", "localsearch", "dlb", "nnls"))
```

This setting is useful to visualize the association between parameters that, for example,
seem to interact. For configuration scenarios that define a large number of parameters, 
it is impossible to visualize all parameters in one of these plots. If do not know which 
parameters to select, all parameters can be split in different plots using the `by_n_param` 
argument which specifies the maximum number of parameters to include in a plot. The functions 
will generate as many plots as needed to cover all parameters included.

## Sampled values and frecuencies

The `sampling_pie` function creates a plot that displays the values of all configurations
sampled during the configuration process. This plot can be useful to display the tendencies
in the sampling in a simple format. As well as in the `previous`parallel_cat` plot, numerical 
parameters domains are discretized to be shown in the plot. The size of each parameter
value in the plot is dependent of the number of configurations having that value in the
configurations.

```{r fig.align="center", fig.width= 7, message=FALSE, prompt=FALSE}
sampling_pie(irace_results = iraceResults)
```

As in previous plots, you can select the parameters to display with the `param_names`
argument and the number of intervals of each parameter with the `n_bins` argument. The
generated plot is interactive, you can click in each parameter to display it 
independently.

In some cases in might be interesting to have a look at the values sampled during the 
configuration procedure as a distribution. Such plot shows the areas in the parameter 
space where irace detected a high performance. A general overview of the distribution of
sampled parameters values can be obtained with the `sampling_frequency`function which 
generates frequency and density plots for the sampled values:

```{r fig.align="center", fig.width= 7, fig.height=7, message=FALSE, prompt=FALSE}
 sampling_frequency(iraceResults)
```

If you would like to visualize the distribution of a particular set of configurations, 
you can pass directly a set of configurations 
and a parameters object in the irace format to the `sampling_frequency` function:
```{r fig.align="center", fig.width= 7, fig.height=7, message=FALSE, prompt=FALSE, eval=FALSE}
 sampling_frequency(iraceResults$allConfigurations, iraceResults$parameters)
```

The previous functions display the parameter frequency plots grouped by 9 plots, you can
adjust this setting using the `n` argument. You can select the parameters to be displayed,
with the `param_names` argument. You can use these plots to ge a general idea of the area 
of the parameter space in which the sampling performed by irace was focused. For example:

```{r fig.align="center", fig.width= 7, message=FALSE, prompt=FALSE}
 sampling_frequency(iraceResults, param_names = c("beta"))
```

If you would like to see more details, a plot showing the sampling by iteration can be 
obtained with the `sampling_frequency_iteration`function. This plot shows the convergence 
of the configuration process reflected in the parameter values sampled each iteration:

```{r fig.align="center", fig.width= 7, fig.height=6}
sampling_frequency_iteration(iraceResults, param_name = "beta")
```

In some cases, you may want to assess how the sampling frequencies of two parameters are 
related. You can visualize the joint sampling frequency of two parameters using the 
`sampling_heatmap` function. By default, this plot uses the elite configurations values 
(of all configurations), you can allow the plot to show all sampled configurations using 
the `only_elite` argument. You must select two parameters using the `param_names` argument, 
for example:


```{r fig.align="center", fig.width= 7, fig.height=6}
sampling_heatmap(iraceResults, param_names = c("beta","alpha"))
```

You can also select the iterations from which configurations will be selected using
the `iterations` argument. The size of the intervals considered for numerical parameters 
in the heatmap can be also adjusted, see the example to have more details.

If you would like to display a set of configurations directly provided by you, 
use the `sampling_heatmap2` function. In both `sampling_heatmap2` and `sampling_heatmap2`, 
you can adjust the number of intervals to be displayed for numerical parameters
using the `sizes` argument. For example, we set to 5 the number of intervals displayed 
for the second parameter, which is numerical, using `sizes=c(0,5)`. The the 0 value in 
this argument indicates that the default interval size should be used. In this case, we set 
0 for the first parameter but note that it is not possible to adjust the number of intervals 
for categorical or ordered parameter types.

```{r fig.align="center", fig.width= 7, fig.height=6}
sampling_heatmap2(iraceResults$allConfigurations, iraceResults$parameters, 
                  param_names = c("localsearch","nnls"), sizes=c(0,5))
```

## Sampling distance 

You may like to have a general overview of the distance of the configurations sampled
across the configuration process. This can allow you to assess the convergence of the
configuration process. The mean distance between the sampled configurations can be 
visualized using the `sampling_distance` function. This function compares the parameter
values of all configurations and aggregates these comparisons to calculate the overall 
distance:

```{r fig.align="center", fig.width= 7}
sampling_distance(iraceResults, t=0.05)
```

Note that for categorical and ordered parameters the comparison is straightforward, but 
this is not the case for numerical parameters. The argument `t` defines a percentage 
used to define a domain interval to assess equality of numerical parameters. For example, 
if the domain of a parameter is `[0,10]` and `t=0.1` then when comparing any value to `v=2` 
we define an interval `s=t* (upper_bound -lower_bound) = 0.1*(10-0)=1`. Then all 
values in the interval `[v-s, v+s] [1,3]` will be equal to `v=2`.

# Visualizing Performance

## Test performance (elite configurations)

When executing irace, you can enable the testing feature that will evaluate elite configurations
in a set of test instances. For more details about how you can use this feature check the irace package 
[user guide](https://cran.r-project.org/web/packages/irace/vignettes/irace-package.pdf).

The test performance of the evaluated configurations can be visualised using the `boxplot_test`
function. Note that the irace execution log file includes test data (test is not enabled by default).

```{r fig.align="center", fig.width=7}
boxplot_test(iraceResults, type="best")
```

This plot shows all final elite configurations evaluation on the test instance set, we can compare
the performance of these configurations to select one that has the best test performance. Note that 
the best configuration (identified by irace) is displayed in a different color. By default, the plot
displayes the relative percentage deviation of the performance of the configurations, to disable this
and display the raw performance use the `rpd=FALSE` argument.

If the irace log file includes the evaluation on the test set of the iteration elite configurations, 
its possible to plot the test performance of elite configurations across the iterations using the 
argument `type="all"`. The best elite configuration on each iteration is displayed in different color:

```{r fig.align="center", fig.width=7}
boxplot_test(iraceResults, type="all", show_points=FALSE)
```

This plot allows to assess the progress of the configuration process regarding the test set performance, 
which would be useful when dealing with heterogeneous instance sets. In these cases, good configurations 
across the full set can be challenging to find and it is possible that the algorithm could be mislead if 
instances sets are prone to introduce bias due to instance ordering. 

Note that in this example the elite configuration with id 808 seems to have a slightly better performance  
than the configuration identified as the best (id:809) by irace. It is important to note that all elite
configurations are not statistically different and thus, its very possible that such situation is observed
when evaluuting test performance, specially when configuring heterogeneous, difficult to balance, instance 
sets. 

If you would like further detauls about the difference in the performance of two configurations, 
you can use the `scatter_test` function. This function displays the performance of both configurations 
paired by instance (each point represents an instance):

```{r fig.align="center", fig.width=7}
scatter_test(iraceResults, x_id = 808, y_id = 809, interactive=TRUE, instance_names = basename)
```

If the plot is created using the argument `interactive=TRUE` you can visualize the instance name when
placing the cursor over each performance point. This plot can help to identify subsets of instances 
in which a configuration clearly outperforms other. To further understand the difference of these two
configurations the trainig data might be explored to verify if such effect holds for the training set.


## Training performance (all configurations)

During the execution of irace considerable performance data is obtained in order to 
assess the performance of the candidate configurations. This data can be useful to
understand the performance of the configured algorithm and to get insights about how
to improve the configuration process.

The following functions create plots of the training data in the irace log. Note that this 
data is obtained during the search of good configurations. Due to the racing procedure, 
some configurations are more evaluated than others (best configurations are more evaluated 
than poor performing configurations). See the irace package documentation for details.

Visualizing training performance might help to obtain insights about the reasoning
that followed irace when searching the parameter space, and thus it can be used
to understand why irace considers certain configurations as high or low  performing.

To visualize the performance of the final elites as observed by irace, the `boxplot_training`
function plots the experiments performed on these configurations. Note that this data corresponds
to the performance generated during the configuration process thus, the number of instances on
which the configurations were evaluated might vary between elite configurations.


```{r fig.align="center", fig.width=7}
boxplot_training(iraceResults)
```

You can select to display the elite configurations of a different iteration by using the 
`iteration`. In case you would like to visualize non-elite configurations you can 
directly provide the instance ids using the `. This can be very useful to assess the performance of 
initial configurations provided for the configuration process for example, providing 
insights about why irace did not select them as final elite configurations:

```{r fig.align="center", fig.width=7}
iteration_elites = iraceResults$iterationElites
boxplot_training(iraceResults, id_configurations=c(1, iteration_elites))
```

To visualize the difference in the performance of two configurations you can also generate
a scatter plot using the `scatter_training` function: 

```{r fig.align="center", fig.width=7}
scatter_training(iraceResults, x_id = 808, y_id = 809, interactive=TRUE)
```

If the plot is created using the argument `interactive=TRUE` you can visualize the instance name when
placing the cursor over each performance point. 

For both functions `boxplot_training` and `scatter_training`, you can display either the 
relative percentage deviation (`rpd=TRUE`) or the raw performance (`rpd=FALSE`). 

You can also plot the performance of configurations which was not necessarily obtained 
when executing irace. Check the General purpose performance section for details. 


## General purpose performance 

You can use the following functions to plot the performance of a selected set 
of configurations in an experiment matrix provided directly by you. These functions 
can be also useful when you would like to compare a configuration that 
was not generated in the configuration process. Note thar such comparison should be
carefully considered as execution conditions might differ from the ones when irace was 
executed. Keep in mind that irace provides seeds to execute instances when configuring 
stochastic algorithms and also might define an execution bound when adaptive capping is 
active. Check the irace package [user guide](https://cran.r-project.org/web/packages/irace/vignettes/irace-package.pdf) 
for details about this.

To plot the performance of a selected set of configurations in an experiment matrix, 
you can use the `boxplot_performance` function. The configurations can be selected
in a vector (allElites):

```{r fig.align="center", fig.width=7}
boxplot_performance(iraceResults$experiments, allElites=c(800, 803,808,809))
```

The experiment matrix should be provided in the irace format, columns should
have configuration ids as names (character) and rows names can be instances 
names (optional).

If the configurations are provided in a list, then the different elements of this list 
will be considered as iterations. Note that this matches the `iraceResults$allElites` 
variable in the irace log. For example, to plot 3 configurations as they are assigned
to 3 different iterations:

```{r fig.align="center", fig.width=7}
boxplot_performance(iraceResults$experiments, allElites=as.list(c(800,803,808,809)))
```

Each element of a list can have more than one id (vector). You can place at the start of 
each  vector the configuration id you want to be identified as the best one and use 
`first_is_best = TRUE` to have it displayed in the different color. Adjust the color using
the `best_color` argument.

```{r fig.align="center", fig.width=7}

boxplot_performance(iraceResults$experiments, allElites=list(c(803,808), c(809,800)), first_is_best = TRUE)
```

If you want to further compare the performance of two configurtions, you can use the 
`scatter_perfomance` function to plot the difference between configurations:

```{r fig.align="center", fig.width=7}
scatter_performance(iraceResults$experiments, x_id = 803, y_id = 809, interactive=TRUE, instance_names = basename)
```

If the plot is created using the argument `interactive=TRUE` and, the provided
matrix has row names, you can visualize the instance name when placing the cursor 
over each performance point otherwise an instance ID is displayed. In this
example, we further transform instance names using the function `basename()`. 

# Visualizing the configuration process

In some cases, it might be interesting have a general visualization for the 
configuration process progress. This can be generated with the `plot_experiments_matrix` 
function:

```{r fig.align="center", fig.width=7}
plot_experiments_matrix(iraceResults, interactive = TRUE)
```

This plot shows configurations in the x axis and instances in the y axis. Each point
in the plot display in color the ranking of the configuration on each instance. If 
`interactive = TRUE`, you can place your cursor on each point to visualize the configuration
id, instance id, and the rank of the configuration.

The sampling distributions used by irace during the configuration process can be displayed using the 
`plot_model` function. For categorical parameters, this function displays the sampling probabilities 
associated to each parameter value by iteration (x axis top) in each elite configuration model (bars):

```{r fig.align="center", fig.width=7}
plot_model(iraceResults, param_name="algorithm")
```

For numerical parameters, this function shows the sampling distributions associated to each parameter.
These plots display the the density function of the truncated normal distribution associated to the 
models of each elite configuration in each instance:


```{r fig.align="center", fig.width=7, fig.height=6, message=FALSE, prompt=FALSE, results='hide'}
plot_model(iraceResults, param_name="alpha")
```

# Report

If you want a quick and portable overview of the configuration process, you can use 
the `report` function which generates an HTML report with a summary of the configuration 
process executed by irace. The function will create an HTML file in the path
provided in the `filename` argument and appending the `".html"` extension to it.


```{r fig.align="center", eval=FALSE}
report(iraceResults, filename="report")
```

