---
title: "Report Iraceplot Package"
output: pdf_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include= FALSE}
library(iraceplot)
```

## Testing performance box plot

  The ``boxplot_test`` function creates a box plot using the performance of the configurations tested. Note that the performance that can be displayed is restricted to the testing performed by irace.
  
```{r,fig.align="center", fig.width= 7}
boxplot_test(iraceResults, type = "best")
```

```{r,fig.align="center", fig.width= 7}
boxplot_test(iraceResults, type = "ibest")
```

## Training performance box plot

The ``boxplot_training`` function creates a box plot using the performance of the configurations obtained in the configuration process. Note that this function can only plot the performance of elite configurations.

```{r,fig.align="center", fig.width= 7}
boxplot_training(iraceResults)
```

## Parallel coordinates 

The ``parallel_coord`` function creates a parallel coordinate plot of a set of selected configurations. 

# ```{r,fig.align="center", fig.width= 7}
# parallel_coord(iraceResults)
# ```

## Sampling proportions

The ``sampling_pie`` function creates a pie chart to visualize the proportion of the sampling values for categorical parameters in the configuration process.


# ```{r,fig.align="center", fig.width= 7}
# sampling_pie(iraceResults)
# ```

## Sampling frequency

The ``sampling_frequency`` function plots the sampling frequency plots of the parameter values sampled during the configuration process.

For categorical parameters frequency bar plot is displayed, while for of numerical parameters (integer and real) an histogram and density plot are shown.


```{r,fig.align="center", fig.width= 7}
sampling_frequency(iraceResults)
```

